### 2012-09-20 doc/prelim/document/chapters/linear_problem/linear_problem.tex
adding reproducability section

\subsubsection{Domain Decomposition for Monte Carlo}
\label{subsubsec:msod}
As observed in the discussion on parallel Krylov methods, large-scale
problems will surely have their data partitioned such that each
parllel process owns a subset of the equations in the linear
system. Given this convention, the adjoint Monte Carlo algorithm must
perform random walks over a domain that is decomposed and must remain
decomposed due to memory limitations. This naturally leads us to seek
parallel algorithms that handle domain decomposition. In the context
of radiation transport, Brunner and colleagues provided a survey of
algorithms for achieving this as implemented in production implicit
Monte Carlo codes \citep{brunner_comparison_2006}. In their work they
identity two data sets that are required to be communicated: the
sharing of particles that are transported from one domain to another
and therefore from one processor to another and a global communication
that signals if particle transport has been completed on all
processors. The algorithms presented are a fully-locking synchronous
scheme, a asynchronous-send/synchronous-receive pattern, a traditional
master/slave scheme, and a modified master/slave scheme the implements
a binary tree pattern for the global reduction type operations needed
to communicate between the master and slave processes. They observed
that the modified master/slave scheme performed best in that global
communications were implemented more efficiently than those required
by the asynchronous scheme. Furthermore, none of these schemes handled
load-imbalanced cases efficiently. Such cases will be common if the
source sampled in the Monte Carlo random walk is not isotropic and
evenly distributed throughout the global domain. It was noted that
efficiencies were improved by increasing the frequency by which
particle data was communicated between domain-adjacent processors.
However, this ultimately increases communication costs.

The 2006 work of Brunner is notable in that the Monte Carlo codes used
to implement and test the algorithms adhered to a strict policy of
generated identical results, indepdent of domain decomposition or
domain replication as derived from the work of Gentile and colleagues
\citep{gentile_obtaining_2005}. In their work, a procedure is given for
obtaining results reproducable to machine precision for an arbitrary
number of processors and domains. Differences can arise from using a 
different random number sequence in each domain and performing
a sequence of floating point operations on identical data in a
different, leading to variations in round-off error and ultimately a
non-identical answer. They use a simple example, recreated below in
Figure~\ref{fig:gentile_example}, that illustrates these issues.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/gentile_example.pdftex_t} }
  \end{center}
  \caption{\textbf{Gentile's example illustrating how domain
      decomposition can create reproducability issues in Monte Carlo.}
    \textit{Both particles A and B start in zone 1 on processor
      1. Particle A moves to zone 2 on processor 2 and scatters back
      to zone 1 while B scatters in zone 1 and remains there.}}
  \label{fig:gentile_example}
\end{figure}

@@

>>>>

\subsubsection{Domain Decomposition for Monte Carlo}
\label{subsubsec:msod}
As observed in the discussion on parallel Krylov methods, large-scale
problems will surely have their data partitioned such that each
parllel process owns a subset of the equations in the linear
system. Given this convention, the adjoint Monte Carlo algorithm must
perform random walks over a domain that is decomposed and must remain
decomposed due to memory limitations. This naturally leads us to seek
parallel algorithms that handle domain decomposition.

In the context of radiation transport, Brunner and colleagues provided
a survey of algorithms for achieving this as implemented in production
implicit Monte Carlo codes \citep{brunner_comparison_2006}. In their
work they identity two data sets that are required to be communicated:
the sharing of particles that are transported from one domain to
another and therefore from one processor to another and a global
communication that signals if particle transport has been completed on
all processors. The algorithms presented are a fully-locking
synchronous scheme, a asynchronous-send/synchronous-receive pattern, a
traditional master/slave scheme, and a modified master/slave scheme
the implements a binary tree pattern for the global reduction type
operations needed to communicate between the master and slave
processes. They observed that the modified master/slave scheme
performed best in that global communications were implemented more
efficiently than those required by the asynchronous
scheme. Furthermore, none of these schemes handled load-imbalanced
cases efficiently. Such cases will be common if the source sampled in
the Monte Carlo random walk is not isotropic and evenly distributed
throughout the global domain. It was noted that efficiencies were
improved by increasing the frequency by which particle data was
communicated between domain-adjacent processors.  However, this
ultimately increases communication costs.

@@

\subsubsection{Reproducable Domain Decomposed Results}
The 2006 work of Brunner is notable in that the Monte Carlo codes used
to implement and test the algorithms adhered to a strict policy of
generating identical results indepdent of domain decomposition or
domain replication as derived from the work of Gentile and colleagues
\citep{gentile_obtaining_2005}. In their work, a procedure is given for
obtaining results reproducable to machine precision for an arbitrary
number of processors and domains. Differences can arise from using a 
different random number sequence in each domain and performing
a sequence of floating point operations on identical data in a
different, leading to variations in round-off error and ultimately a
non-identical answer. They use a simple example, recreated below in
Figure~\ref{fig:gentile_example}, that illustrates these issues.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/gentile_example.pdftex_t} }
  \end{center}
  \caption{\textbf{Gentile's example illustrating how domain
      decomposition can create reproducability issues in Monte Carlo.}
    \textit{Both particles A and B start in zone 1 on processor
      1. Particle A moves to zone 2 on processor 2 and scatters back
      to zone 1 while B scatters in zone 1 and remains there. A1 and
      A2 denote the track of particle A that is in zone 1 while B1 and
      B2 denote the track of particle B that is in zone 1.}}
  \label{fig:gentile_example}
\end{figure}
In this example, the domain is decomposed on two processors with each
processor owning one of the two zones. Starting with particle A, it is
born in zone 1 and is transported to zone 2 where a scattering event
occurs. Concerning the first reproducability issue, if the same
sequence of random numbers is not used to compute the trajectory from
the new scattering event, we cannot expect to achieve the same result
if the domain were not decomposed. If the numbers are different, the
scattering event in zone 2 may keep the particle there or even eject
it from the domain if the sequence were different. The second issue is
demonstrated by adding another particle B that remains in the
domain. In this case, an efficient algorithm will transport particle A
on processor 1 until it leaves zone 1 and then transport particle
B. Particle A will not renter the domain until it has been
communicated to processor 2, processor 2 performs the transport, and
it is communicated back to processor 1. If we are doing a track-length
tally in zone 1, then we sum the tracks lengths observed in that
zone. In the single processor, single zone case particle A would be
transported in its entirety and then particle B transported. This
would result in a tally sum with the following order of operations:
$(((A1+A2)+B1)+B2)$. If we were instead to use 2 processors, we would
instead have the following order: $(((A1+B1)+B2)+A2)$. In the context
of floating point operations, we cannot expect these to have an
identical result to machine precesion as round-off approximations will
differ.
